Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-05-19 manuscript/response_to_reviews.Rmd

The study addresses the relationship between drought resistance (Rt) calculated from radial growth, and different traits and microenvironmental characteristics on 12 tree species from one forest in the Eastern US. Both the dataset and the study are meaningful, and the relationship found between Rt and height of biological importance. However, I think there are a number of issues that need to be worked out.

It is important to better discuss the proper use of Rt, particularly when comparing single drought years, to assess tree resistance to drought. This metric has been much used lately as one of the so-called resilience indices (I wonder why other Rs indices where not included in the manuscript). In the paper, the authors use Rt similarly to most studies in the literature, which is fine. Yet, Rt (and particularly the other resilience indices based on radial growth) does not only depend on the individual response to drought events, but also on stand dynamics around target trees. For this reason, if stands around trees are not comparable or show changes after some disturbance (e.g. the drought events analysed) differently, this may bias comparisons of Resilience indices (although it is true that Rt should be less influenced than the other indices). The problem is that there is not such stand competition data for old droughts. This does not invalidate the study but I think the authors should discuss the issue, particularly for the analysis and comparisons of single droughts. I would suggest including also other Resilience indices, not only Rt, to enrich the discussion.

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

**Regarding the resilience indices, we considered analyzing these and decided it would be best to stick with just resistance for three reasons: (1) Most importantly, we would expect resilience to be governed by fundamentally different mechanisms than resistance. While it would be interesting to examine resilience, this would be beyond the scope of the current paper. (2) [DeSoto et al. (2020, Nature)](https://www.nature.com/articles/s41467-020-14300-5) used an extensive dataset of trees sampled after severe droughts and compared the responses to previous, less-severe droughts between trees that survived and trees that were killed by the severe drought. In angiosperms, trees that died from a severe drought had lower resistance to previous droughts, compared to trees that survived the severe drought. In gymnosperms, trees that died during severe droughts had lower recovery following previous droughts, compared to trees that survived the severe drought. These findings would argue that resistance is a more important variable for angiosperms. (3) As the reviewer notes, resilience indices would probably tend to be more sensitive to a tree's neighborhood and prior conditions. We now state the logic for focusing on $Rt$ in the methods section:**

**"We focus exclusively on Rt, and not on the resilience metrics described in Lloret et al. (2011), because (1) we would expect resilience to be controlled by a different set of mechanisms, and (2) the findings of DeSoto et al. (2020) suggest that Rt is a more important drought response metric for angiosperms."**

@@

Firstly, the way drought resistance (Rt) is estimated, as the ratio of BAI during drought to the mean BAI over the five years preceding the drought can be problematic, as growth dynamics during this pre-drought period were quite variable across species (Figure 1; incidentally, why do you show ring width indices in this figure if you then use BAI?). This approach has also advantages, particularly when applied in a resilience framework (as in the original Lloret et al. (2011)). However, if you focus specificaly on resistance it seems that directly modelling the growth reduction caused by the drought (e.g., using intervention time series analysis or other methods) would be a more robust approach. At the very least, if you continue using your simpler approach I would strongly suggest including some measure of pre-drought growth in the models to account for differences in growth dynamics among species. It would also nice to clarify why you do not assess recovery and resilience, unlike Lloret et al. (2011).

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

**Regarding the resilience indices, we considered analyzing these and decided it would be best to stick with just resistance for three reasons: (1) Most importantly, we would expect resilience to be governed by fundamentally different mechanisms than resistance. While it would be interesting to examine resilience, this would be beyond the scope of the current paper. (2) [DeSoto et al. (2020, Nature)](https://www.nature.com/articles/s41467-020-14300-5) used an extensive dataset of trees sampled after severe droughts and compared the responses to previous, less-severe droughts between trees that survived and trees that were killed by the severe drought. In angiosperms, trees that died from a severe drought had lower resistance to previous droughts, compared to trees that survived the severe drought. In gymnosperms, trees that died during severe droughts had lower recovery following previous droughts, compared to trees that survived the severe drought. These findings would argue that resistance is a more important variable for angiosperms. We now state the logic for focusing on $Rt$ in the methods section:**

**"We focus exclusively on Rt, and not on the resilience metrics described in Lloret et al. (2011), because (1) we would expect resilience to be controlled by a different set of mechanisms, and (2) the findings of DeSoto et al. (2020) suggest that Rt is a more important drought response metric for angiosperms."**

### 2020-05-19 manuscript/climate_sensitivity_draft.Rmd
response on resilience metrics
#93

For each drought year, we calculated drought resistance ($Rt$) as the ratio of $BAI$ during drought to the mean $BAI$ over the five years preceding the drought [@lloret_components_2011]. Thus, $Rt$ values <1 and >1 indicate growth reductions and increases, respectively. Analyses focused on testing the predictions presented in Table 1, with $Rt$ as the response variable. The general statistical model for hypothesis testing was a mixed effects model with $Rt$ as the response variable, tree nested within species as a random effect, and one or more independent variables as fixed effects. Mixed effects models were implemented in the lme4 package in R [@R-lme4]. We used AICc to assess model selection, and conditional/marginal R-squared to assess model fit, implmented in the AICcmodavg package in R [@R-AICcmodavg].

>>>>

For each drought year, we calculated drought resistance ($Rt$) as the ratio of $BAI$ during drought to the mean $BAI$ over the five years preceding the drought [@lloret_components_2011]. Thus, $Rt$ values <1 and >1 indicate growth reductions and increases, respectively. Analyses focused on testing the predictions presented in Table 1, with $Rt$ as the response variable. We focus exclusively on $Rt$, and not on the resilience metrics described in @lloret_components_2011, because (1) we would expect resilience to be controlled by a different set of mechanisms, and (2) the findings of *[DeSoto et al. (2020)](https://www.nature.com/articles/s41467-020-14300-5)* suggest that $Rt$ is a more important drought response metric for angiosperms. The general statistical model for hypothesis testing was a mixed effects model with $Rt$ as the response variable, tree nested within species as a random effect, and one or more independent variables as fixed effects. Mixed effects models were implemented in the lme4 package in R [@R-lme4]. We used AICc to assess model selection, and conditional/marginal R-squared to assess model fit, implmented in the AICcmodavg package in R [@R-AICcmodavg].
